An open reproducible framework for the study of the iterated prisoner's dilemma

### 2016-03-08 tex/paper.tex
Some tidbits in intro

\begin{abstract}
\noindent
    The Axelrod Python library is an open source package that allows for
    reproducible game theoretic research in to the iterated Prisoner's Dilemma.
    This area of research began in the 1980 but suffers from a lack of
    documented and tested code.

\noindent
    With a growing collection of 120 strategies the library is also a platform
    for an original tournament that in itself is of interest to the game
    theoretic community.

\noindent
    This paper describes the iterated Prisoner's Dilemma, the library and
    insights gained from some novel research.
\end{abstract}

@@

Several iterated prisoner's dilemma tournaments have generated much interest,
including Axelrod's original tournaments, two 2004 anniversary tournaments, and
the Stewart and Plotkin 2012 tournament \cite{Stewart2012} following the
discovery of zero-determinant strategies.  Subsequent research has spawned an
enormous number of papers, but rarely are the results reproducible. Among
well-known tournaments, in only one case is the full original source code
available (Axelrod's second tournament \cite{Axelrod1980b}, in FORTRAN); in no
cases is the available code well-documented, easily modifiable, or released with
significant test suites.

To make matters more complicated, often a newly-created strategy is studied in
isolation by opponents chosen by the strategy's creators, and often such
strategies are not sufficiently described to enable reliable recreation
(in the absense of source code), with \cite{slany2007some} being a notable
counter-example. In some cases strategies are revised without updates to their
names or published implementations \cite{li2007design, li2011engineering}.
As a result, some of the results related to these strategies and tournaments
have not been reliably replicated, and have not met the basic scientific
criterion of falsifiability.

@@

This library is motivated by an ongoing discussion in the academic community
about reproducible research \cite{Crick2014a, Hong2015a, Prlic2012, Sandve2013},
and is:

@@

As stated in~\cite{Bendor1991}: ``\textit{few works in social science have had
the general impact of [Axelrod's study of the evolution of cooperation]}''.  In
1980, Axelrod wrote two papers:~\cite{Axelrod1980a,Axelrod1980b} which
described a computer tournament that has been at the origin of a majority of
game theoretic work~\cite{Banks1990, Bendor1991, Boyd1987, Chellapilla1999,
DavidB1993, Doebeli2005, Ellison1994, Gotts2003, Hilbe2013, Isaac2008,
Kraines1989, Lee2015, Lorberbaum1994, Milgrom1982, Molander1985, Murnighan2015,
Press2012, Stephens2002, Stewart2012}. As described in~\cite{Bendor1991} this
work has not only had mathematical impact but has also led to insights in
biology (for example in~\cite{Stephens2002}, a real tournament where Blue Jays
are the participants is described) and in particular to the study of evolution.

@@

@@

The game of (\ref{equ:one-shot}) is called the Prisoner's Dilemma. Specific
numerical values of \((R,S,T,P)=(3,0,5,1)\) are often used in the literature.
Axelrod's tournaments (and further implementations of these) are sometimes
referred to as Iterated Prisoner's Dilemma (IPD) tournaments. An incomplete
overview of published tournaments is given in Table~\ref{tab:tournaments}.

\begin{table}[!hbtp]
    \begin{center}
        \begin{tabular}{ccccc}
            \toprule
            Year     & Reference                  & Number of Strategies & Type     & Source Code\\
            \midrule
            1979     & \cite{Axelrod1980a}        & 13                   & Standard & Not immediately available\\
            1979     & \cite{Axelrod1980b}        & 64                   & Standard & Available in FORTRAN\\
            1991     & \cite{Bendor1991}          & 13                   & Noisy    & Not immediately available\\
            2002     & \cite{Stephens2002}        & 16                   & Wildlife & Not applicable\\
            2005     & \cite{kendall2007iterated} & 223                  & Varied   & Not available \\
            2012     & \cite{Stewart2012}         & 13                   & Standard & Not fully available \\
            \bottomrule
        \end{tabular}
    \end{center}
    \caption{An overview of published tournaments}\label{tab:tournaments}
\end{table}

@@

Due to the open nature of the library the number of strategies included has
grown at a fast pace, as can be seen in
Figure~\ref{fig:number_of_strategies_against_date}. Despite this, due to
previous research being done in an irreproducible manner, with for example no
source code, and/or vaguely described strategies, not all previous tournaments
are yet to be reproduced. In fact, some of the early tournaments might be
impossible to reproduce as the source code is forever lost, this library aims
to prevent this from happening again in the future.

>>>>

\begin{abstract}
\noindent
    The Axelrod library is an open source Python package that allows for
    reproducible game theoretic research into the iterated Prisoner's Dilemma.
    This area of research began in the 1980s but suffers from a lack of
    documented and test code. The goal of the library is to provide such
    a resource, with facilities the design new strategies and interactions
    between them, as well as conducting tournaments and ecological simulations
    for population of strategies.

\noindent
    With a growing collection of 120 strategies the library is a also a platform
    for an original tournament that in itself is of interest to the game
    theoretic community.

\noindent
    This paper describes the iterated Prisoner's Dilemma, the Axelrod library
    and its development, and insights gained from some novel research.
\end{abstract}

@@

Several iterated prisoner's dilemma tournaments have generated much interest,
including Axelrod's original tournaments \cite{Axelrod1980a,Axelrod1980b},
two 2004 anniversary tournaments (ref?), and
the Stewart and Plotkin 2012 tournament \cite{Stewart2012} following the
discovery of zero-determinant strategies.  Subsequent research has spawned an
enormous number of papers, but rarely are the results reproducible. Among
well-known tournaments, in only one case is the full original source code
available (Axelrod's second tournament \cite{Axelrod1980b}, in FORTRAN); in no
cases is the available code well-documented, easily modifiable, or released with
significant test suites.

To make matters more complicated, often a newly-created strategy is studied in
isolation by opponents chosen by the strategy's creators, and often such
strategies are not sufficiently described to enable reliable recreation
(in the absense of source code), with \cite{slany2007some} being a notable
counter-example. In some cases strategies are revised without updates to their
names or published implementations \cite{li2007design, li2011engineering}.
As a result, some of the results related to these strategies and tournaments
cannot be reliably replicated, and therefore have not met the basic scientific
criterion of falsifiability.

@@

The presented library is partly motivated by an ongoing discussion in the academic community
about reproducible research \cite{Crick2014a, Hong2015a, Prlic2012, Sandve2013},
and is:

@@

As stated in~\cite{Bendor1991}: ``\textit{few works in social science have had
the general impact of [Axelrod's study of the evolution of cooperation]}''.  In
1980, Axelrod wrote two papers:~\cite{Axelrod1980a,Axelrod1980b} which
describe a computer tournament that has been a imajor influence on
subsequent game theoretic work~\cite{Banks1990, Bendor1991, Boyd1987, Chellapilla1999,
DavidB1993, Doebeli2005, Ellison1994, Gotts2003, Hilbe2013, Isaac2008,
Kraines1989, Lee2015, Lorberbaum1994, Milgrom1982, Molander1985, Murnighan2015,
Press2012, Stephens2002, Stewart2012}. As described in~\cite{Bendor1991} this
work has not only had impact in mathematics but has also led to insights in
biology (for example in~\cite{Stephens2002}, a real tournament where Blue Jays
are the participants is described) and in particular in the study of evolution.

@@

and the following reward matrix results from the Cartesian product of
two decision vectors $\langle C, D \rangle$,

@@

The game of (\ref{equ:one-shot}) is called the Prisoner's Dilemma. Specific
numerical values of \((R,S,T,P)=(3,0,5,1)\) are often used in the literature,
although any satisfying the conditions in \ref{equ:one-shot} will yield similar results.
Axelrod's tournaments (and further implementations of these) are sometimes
referred to as Iterated Prisoner's Dilemma (IPD) tournaments. An incomplete
overview of published tournaments is given in Table~\ref{tab:tournaments}.

\begin{table}[!hbtp]
    \begin{center}
        \begin{tabular}{ccccc}
            \toprule
            Year     & Reference                  & Number of Strategies & Type     & Source Code\\
            \midrule
            1979     & \cite{Axelrod1980a}        & 13                   & Standard & Not immediately available\\
            1979     & \cite{Axelrod1980b}        & 64                   & Standard & Available in FORTRAN\\
            1991     & \cite{Bendor1991}          & 13                   & Noisy    & Not immediately available\\
            2002     & \cite{Stephens2002}        & 16                   & Wildlife & Not applicable\\
            2005     & \cite{kendall2007iterated} & 223                  & Varied   & Not available \\
            2012     & \cite{Stewart2012}         & 13                   & Standard & Not fully available \\
            \bottomrule
        \end{tabular}
    \end{center}
    \caption{An overview of published tournaments (specify what "type" means)}\label{tab:tournaments}
\end{table}

@@

Due to the open nature of the library the number of strategies included has
grown at a fast pace, as can be seen in
Figure~\ref{fig:number_of_strategies_against_date}. Despite this, due to
previous research being done in an irreproducible manner, with for example no
source code, and/or vaguely described strategies, not all previous tournaments
are yet to be reproduced. In fact, some of the early tournaments might be
impossible to reproduce as the source code is forever lost, and this library aims
to prevent this from happening again in the future.
